*Bromus* Initial Analysis

### 2019-07-30 Scripts/BromusUpdate30Jul2019.Rmd
Begin new update script for analyses with density and log(lambda)

---
title: '*Bromus* Initial Analysis'
author: "Nick Gotelli"
date: "7/19/2019"
output:
  pdf_document: default
  html_document:
    highlight: tango
    theme: cerulean
---

@@

## Data sources and cleaning
These data were provided by Alden Griffith and Michael Loic. After talking with Alden, we decided to use lambda values that had been estimated from changes in the number of individuals in marked plots, rather than analysis of stage transition matrices. For a number of control and snow-manipulated plots, we have the estimated lambda $(N_{t + 1}/N_{t})$ based on annual counts of cheatgrass for 7 consecutive years.

@@

df_preds <- df_clean[,8:18]           # create dataframe of predictor variables
```

@@

quad_fit <- function(x=df_clean$Snow.Cover, y=df_clean$Lambda){
  M1 <- summary(lm(y~x))
  M2 <- summary(lm(y~poly(x,2)))
  return(c(M1$adj.r.squared,M2$adj.r.squared))
}
# -------------------------------------------------------------------------
# map all predictor variables to quad_fit
summary_df <- map(df_preds,quad_fit,y=df_clean$Lambda)

@@

A couple things jump out from these univariate regressions. First, the quadratic model almost always fits better than the corresponding linear model (even after imposing a penalty for the additional parameter), so we have good theoretical grounds and empirical support for going with the quadratic models.

@@

```{r}
# best variables are:
# Grow.Season.DD.all.days (4)
# Grow.Season.Precip.Total (5)
# Snow.Cover (6)
best_preds <- df_preds[,4:6]

y <- df_clean$Lambda
x_1 <- best_preds[,1]
x_2 <- best_preds[,2]
x_3 <- best_preds[,3]

model_1 <- summary(lm(y~poly(x_1,2) + poly(x_2,2) + poly(x_3,2)))
print(model_1)

@@

lambda_model <- function(y=df_clean$Lambda,z=best_preds){
  y <- df_clean$Lambda
  x_1 <- z[,1] # Grow.Season.DD.all.days
  x_2 <- z[,3]  # Snow.Cover
  forecast_model <- lm(y~poly(x_1,2) + poly(x_2,2))
}

@@

# create parameter data frame with column for output
param_df <- expand.grid(snow_cover=seq(from=snow_low, to=snow_high, length.out=20),
                        degree_days=seq(from=dd_low, to=dd_high,length.out=20))
param_df$lambda <- NA

@@

```{r}
qplot(x=best_preds[,1],y=df_clean$Lambda)

@@

lambda_model2 <- function(y=df_clean$Lambda,z=best_preds){
  y <- df_clean$Lambda
  # x_1 <- z[,1] # Grow.Season.DD.all.days
  x_2 <- z[,3]  # Snow.Cover
  forecast_model <- lm(y~ poly(x_2,2))
}

@@

```{r}
df <- data.frame(Snow.Cover=df_clean$Snow.Cover, Lambda=df_clean$Lambda)
plot_2 <- ggplot(df,aes(x=Snow.Cover,y=Lambda)) +
          geom_point() +
          xlab("Snow Cover") +
          ylab("Observed Lambda")

>>>>

---
title: '*Bromus* Initial Analysis'
author: "Nick Gotelli"
date: "7/30/2019"
output:
  pdf_document: default
  html_document:
    highlight: tango
    theme: cerulean
---

@@

## Data sources and cleaning
These data were provided by Alden Griffith and Michael Loic. After talking with Alden, we decided to use lambda values that had been estimated from changes in the number of individuals in marked plots, rather than analysis of stage transition matrices. For a number of control and snow-manipulated plots, we have the estimated lambda $(N_{t + 1}/N_{t})$ based on annual counts of cheatgrass for 7 consecutive years. For incorporating density dependence, we will use $log(\lambda) = r$.

@@

df_preds <- df_clean[,8:18]           # create dataframe of predictor variables
df_clean$r <- log(df_clean$Lambda)    # create response variable r from log(lambda)
```

@@

quad_fit <- function(x=df_clean$Snow.Cover, y=df_clean$r){
  M1 <- summary(lm(y~x))
  M2 <- summary(lm(y~poly(x,2)))
  return(c(M1$adj.r.squared,M2$adj.r.squared))
}
# -------------------------------------------------------------------------
# map all predictor variables to quad_fit
summary_df <- map(df_preds,quad_fit,y=df_clean$r)

@@

A couple things jump out from these univariate regressions. First, the quadratic model often fits better than the corresponding linear model (even after imposing a penalty for the additional parameter), so we have good theoretical grounds and empirical support for going with the quadratic models.

The four strongest predictors were:

- snow cover (quadratic r^2^ = 0.838)
- initial density (linear r^2^ = 0.638)
- winter precip total (quadratic r^2^ = 0.661)
- growing season precipitation rain (quadratic r^2^ = 0.559)

I had originally intended to include one of the temperature variables, but they perform poorly compared to precip and snow cover variables, so we will start with just these initial 4. The linear regression of r versus initial density is the relationship for a simple logistic growth model. The performance of the quadratic model was virtually identical, so there is no evidence for Allee effects or other more complex functions relating per capita growth rate to density. So we will stick with the linear model.

@@

```{r}
str(df_preds)
# best variables are:
# snow.cover (6)
# density (1)
# winter precip total. (11) 
# grow season precip rain (3)

best_preds <- df_preds[,c(6,1,11,3)]

y <- df_clean$r
x_1 <- best_preds[,1]
x_2 <- best_preds[,2]
x_3 <- best_preds[,3]
x_4 <- best_preds[,4]

final_df <- data.frame(r=y,
                       snow_cover=x_1,
                       density=x_2,
                       win_precip <- x_3,
                       gs_rain <- x_4)

model_dd <- summary(lm(data=final_df,r~density))
print(model_dd)
model_dd <- summary(lm(data=final_df,r~poly(density,2)))
print(model_dd)
ggplot(data=final_df,aes(x=density,y=r)) +
  geom_point() +
  stat_smooth(method="lm", se=TRUE, fill=NA,
                formula=y ~ poly(x, 2, raw=TRUE),colour="red") +
  stat_smooth(method="lm", se=TRUE, fill=NA,
                formula=y ~ poly(x, 1, raw=TRUE),colour="blue")

model_1 <- summary(lm(y~poly(x_1,2) + x_2 + poly(x_3,2) + poly(x_4,2)))
print(model_1)

@@

lambda_model <- function(y=df_clean$r,z=best_preds){
  y <- df_clean$r
  x_1 <- z[,1] # Grow.Season.DD.all.days
  x_2 <- z[,3]  # Snow.Cover
  forecast_model <- lm(y~poly(x_1,2) + poly(x_2,2))
}

@@

# create parameter data frame with column for output
param_df <- expand.grid(snow_cover=seq(from=snow_low, to=snow_high, length.out=20),
                        degree_days=seq(from=dd_low, to=dd_high,length.out=20))
param_df$r <- NA

@@

```{r}
qplot(x=best_preds[,1],y=df_clean$r)

@@

lambda_model2 <- function(y=df_clean$r,z=best_preds){
  y <- df_clean$r
  # x_1 <- z[,1] # Grow.Season.DD.all.days
  x_2 <- z[,3]  # Snow.Cover
  forecast_model <- lm(y~ poly(x_2,2))
}

@@

```{r}
df <- data.frame(Snow.Cover=df_clean$Snow.Cover, Lambda=df_clean$r)
plot_2 <- ggplot(df,aes(x=Snow.Cover,y=Lambda)) +
          geom_point() +
          xlab("Snow Cover") +
          ylab("Observed Lambda")
